---
title: "pokemon data demo"
author: "Deyvis Mejia Zambrana"
date: "`r Sys.Date()`"
output: html_document
---
 Tutorial followed and modified from: 
 https://brunomioto.com/posts/intro_ggplot2/en/#the-data
 
 Check out my rpubs site for other scripts: https://rpubs.com/deyvis305/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Packages required for this tutorial
```{r packages}
library(dplyr)
library(ggplot2)
library(ggpath)                     
library(magick)
library(pokemon)
```




# Pokemon data from the "pokemon" package
```{r pokmeon data}
#load dataset into environment
pokemon.df=pokemon::pokemon # '=' or '<-' are assignment operators
```




# Base R plot of pokemon weight and height
```{r base plot}
#base R plotting
plot(pokemon.df$weight, pokemon.df$height)
```

Label the axes. 
```{r}
plot(pokemon.df$weight, pokemon.df$height, 
     xlab ="kg", ylab = "m", main="pokemon heights and weights")
?plot() #use to see documentation or use help(plot)
```

Look at the data.
```{r}
#glimpse the data
glimpse(pokemon.df) #notice the type of each column, any missing values

#head of the data (first rows)
head(pokemon.df)

#tail of the data (last rows)
tail(pokemon.df) #there are a total of 10,147
```

Is there a relationship between height and weight?
```{r}
?cor() #correlation function
cor(pokemon.df$weight, pokemon.df$height, 
    method="pearson")

#okay now add a linear regression line to the previous plot
plot(pokemon.df$weight, pokemon.df$height,
     xlab = "height (m)", ylab="weight (kg)", main="pokemon heights and weights")
abline(lm(height ~ weight,data = pokemon.df),
       col="red")
```


Is there a relationship between weight and speed?
```{r}
#cor
cor(pokemon.df$weight, pokemon.df$speed) #no correlation

#assume that pokemon speeds are in m/s units
plot(x=pokemon.df$weight, y=pokemon.df$speed,
     xlab="weight (kg)",ylab="speed (m/s)", 
     main = "relationship between pokemon speed and weight")
abline(lm(speed ~ weight, data=pokemon.df),
       col="blue")
```




# Using the ggplot package

```{r ggplot}
#ggplot2 is part of the tidyverse packages
#and can use different syntax to do same things
#The + sign means 'add this'
ggplot()                                       #1 graphics layer
ggplot(data=pokemon.df,aes(x=weight,y=height)) #2 graphics layers
ggplot(data=pokemon.df,aes(x=weight,y=height))+
  geom_point()                                 #3 graphics layers
```

Now let's pimp up the ggplot by adding color and shapes.
```{r}
#pimp up the plot! make color relative to type_1 data categories
ggplot(
  data=pokemon.df, aes(x=weight,y=height))+
  geom_point(shape=25,
             aes(color=type_1))+
  ggtitle("heights and weights of pokemon according to type_1")+
  theme_bw()+
  geom_smooth() #adds a smooth regression line using a default method
```

To summartize: there seems to be a maximum for heights as weight increase.
Most of the pokemon are clustered between height 0 to 5 m and 0 to 500 kg.

We can calculate the range, average, and median for the heights and weights
```{r}
#descriptive statistics
range(pokemon.df$height) #heights range
mean(pokemon.df$height)  #average height
median(pokemon.df$height)#median height

range(pokemon.df$weight) #weights range
mean(pokemon.df$weight)  #average weight
median(pokemon.df$weight)#median weight
```
## Box plots
```{r ggplot boxplots}
#box plots of heights for all pokemon FIRST type
ggplot(data=pokemon.df,aes(x=type_1,y=height))+
  geom_boxplot(aes(color=type_1))+
  theme(axis.text.x = element_text(angle=50, vjust=0.5))+
  ggtitle("height distributions based on pokemon's first type")
```
```{r}
#box plots of weights for all pokemon FIRST type
ggplot(data=pokemon.df,aes(x=type_1, y=weight))+
  geom_boxplot(aes(color=type_1))+
  theme(axis.text.x = element_text(angle=50, vjust=0.5))+
  ggtitle("weight distributions based on pokemon's first type")
```

```{r}
#box plots of HP for all pokemon FIRST type
ggplot(data=pokemon.df,aes(x=type_1, y=hp))+
  geom_boxplot(aes(color=type_1))+
  theme(axis.text.x = element_text(angle=50, vjust=0.5))+
  ggtitle("health point distributions based on pokemon's first type")
```

```{r}
#box plots of speeds for all pokemon FIRST type
ggplot(data=pokemon.df,aes(x=type_1, y=speed))+
  geom_boxplot(aes(color=type_1))+
  theme(axis.text.x = element_text(angle=50, vjust=0.5))+
  ggtitle("speed distributions based on pokemon's first type")
```

## Column plots using ggplot

```{r column plot defense}
#column plots of the defense of all pokemons in the dataset
ggplot(data=pokemon.df, aes(x=pokemon,y=defense))+
  geom_col()+
  ggtitle("defense scores for all pokemon")
```


What is the difference between geom_col and geom_bar? 
Use ?geom_col and ?geom_bar.

THE END
